---
title: PTPN11 mutagenesis

params: 
  prefix: "mut"
  gene_file: "../data-raw/gene.fasta"
molstar: embed

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,max.print=10,fig.path = "analysis/plots/",fig.dim=c(12,14))
```


# PTPN11 ***Se***quence ***Sa***turation ***M***utagenesis (SeSaM)

## Intro 

PTPN11 gene is involved in a rare leukemia. The aim of the project is to investigate the effect of functional mutations in different aminoacids of this protein into developing the illness.

The technique introduces mutations by adding *universal nucleoside analogs* in a pcr amplification process. For each copy, these nucleosides are inserted in one single random place across the sequence. Then an elongation process takes place where 0 to 2 extra analogs are introduced. Finally, these analog bases are substituted by random nucleotides. In these project in particular the targeted protein is *PTPN11* and has a sequence of *594 codons* / (593 aa + stop) and it's equivalent CDS of *1782 bp*. 
Let's have a look at the raw gene sequenece:

```{r}
library(data.table)
library(stringi)
wt_seq<- paste0(readLines(params$gene_file)[-1],collapse="")
cat("\nPTPN11 gene sequence:\n\n",wt_seq)

```

```{r}
wt_counts<- data.table(t(stri_count_fixed(wt_seq,c("A","C","T","G"))/nchar(wt_seq) * 100))
colnames(wt_counts)<-c("A","C","T","G")

wt_codons<-as.character(stri_extract_all_regex(wt_seq,'.{1,3}',simplify=T)) #wt_codons <- stri_sub(wt_seq,length=3,from=c(seq(from=1,to=nchar(wt_seq),by=3)))

```

And the "wild type" protein sequence:

```{r}
wt_counts[,(names(table(wt_codons))):= as.list(table(wt_codons) / length(wt_codons)*100)]
ct<-as.data.table(RGenetics::geneticCodeTable(),key = "GeneticCode")
wt_aa <- ct[wt_codons,AminoAcids]
cat("\nPTPN11 amino acid sequence:\n\n",wt_aa)
wt_counts[,(names(table(wt_aa))):= as.list(table(wt_aa) / length(wt_aa)*100)]

```

So, it is quite difficult to get anything from those. Let's take a look into its 3D structure:

{{< mol-rcsb 3B7O >}}

We can also take a look in the [AlphaFold Database](https://alphafold.com/) to compare the wild type structure to the mutants.
{{< mol-afdb Q96I26 >}}
 

